---
title: "p8105_hw2_nr2774"
author: "Nergui"
date: "2022-10-04"
output: github_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, message = FALSE)
```

### Problem 0

This solution focuses on a reproducible report containing code and text necessary for Problems 1-3, and is organized as an R Project. This was not prepared as a GitHub repo; examples for repository structure and git commits should be familiar from other elements of the course.

Throughout, we use appropriate text to describe our code and results, and use clear styling to ensure code is readable. 

```{r load_libraries}
library(tidyverse)
library(readxl)
```


### Problem 1

Below we import and clean data from `NYC_Transit_Subway_Entrance_And_Exit_Data.csv`. The process begins with data import, updates variable names, and selects the columns that will be used in later parts fo this problem. We update `entry` from `yes` / `no` to a logical variable. As part of data import, we specify that `Route` columns 8-11 should be character for consistency with 1-7.

```{r}
#??? entry logical how 
trans_ent = 
  read_csv(
    "hw2_data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv",
    col_types = cols(Route8 = "c", Route9 = "c", Route10 = "c", Route11 = "c")) %>% 
  janitor::clean_names() %>% 
  select(
    line, station_name, station_latitude, station_longitude, 
    starts_with("route"), entry, exit_only, vending, entrance_type, 
    ada) %>% 
  mutate(entry = ifelse(entry == "YES", TRUE, FALSE))
```

As it stands, these data are not "tidy": route number should be a variable, as should route. That is, to obtain a tidy dataset we would need to convert `route` variables from wide to long format. This will be useful when focusing on specific routes, but may not be necessary when considering questions that focus on station-level variables. 

The following code chunk selects station name and line, and then uses `distinct()` to obtain all unique combinations. As a result, the number of rows in this dataset is the number of unique stations.

```{r}
trans_ent %>% 
  select(station_name, line) %>% 
  distinct
```

The next code chunk is similar, but filters according to ADA compliance as an initial step. This produces a dataframe in which the number of rows is the number of ADA compliant stations. 

```{r}
trans_ent %>% 
  filter(ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```

To compute the proportion of station entrances / exits without vending allow entrance, we first exclude station entrances that do not allow vending. Then, we focus on the `entry` variable -- this logical, so taking the mean will produce the desired proportion (recall that R will coerce logical to numeric in cases like this).

```{r}
trans_ent %>% 
  filter(vending == "NO") %>% 
  pull(entry) %>% 
  mean
```

Lastly, we write a code chunk to identify stations that serve the A train, and to assess how many of these are ADA compliant. As a first step, we tidy the data as alluded to previously; that is, we convert `route` from wide to long format. After this step, we can use tools from previous parts of the question (filtering to focus on the A train, and on ADA compliance; selecting and using `distinct` to obtain dataframes with the required stations in rows).

```{r}
trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A") %>% 
  select(station_name, line) %>% 
  distinct

trans_ent %>% 
  pivot_longer(
    route1:route11,
    names_to = "route_num",
    values_to = "route") %>% 
  filter(route == "A", ada == TRUE) %>% 
  select(station_name, line) %>% 
  distinct
```



### Problem 2

```{r, warning = FALSE,message=FALSE}
# Tidy the Mr. Trash Wheel dataset

mr_df = read_excel("./hw2_data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                                 sheet = "Mr. Trash Wheel",
                                 range = cellranger::cell_cols("A:N")) |>  
  janitor::clean_names() |> 
  drop_na(dumpster) |>  
  filter(dumpster != "Grand Total") |> 
  mutate(sports_balls = as.integer(round(sports_balls, 0)),
         month = recode(month,"Decemeber" = "December"),
         dumpster = as.integer(dumpster)) |>  
  rename(weight = weight_tons, volume = volume_cubic_yards)
```



```{r, warning = FALSE,message=FALSE}
# Tidy the Professor Trash Wheel dataset

prof_df = read_excel("./hw2_data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", 
                                 sheet = "Professor Trash Wheel",
                                 range = cellranger::cell_cols("A:N")) |>  
  janitor::clean_names() |> 
  drop_na(dumpster) |>  
  filter(dumpster != "Grand Total") |> 
  mutate(sports_balls = as.integer(round(sports_balls, 0))) |>  
  rename(weight = weight_tons, volume = volume_cubic_yards)
```



```{r, warning = FALSE,message=FALSE}
# Merge the two Trash Wheel datasets
# ???? Why mutate names = Mr. Trash Wheel 

mr_merge_df = mr_df |>  mutate(name= "Mr. Trash Wheel")
prof_merge_df = prof_df |>  mutate(name = "Professor Trash Wheel")
merge_df = bind_rows(mr_merge_df, prof_merge_df)

# Total weight of trash collected by Professor Trash Wheel
total_weight_prof = sum(prof_df$weight)
# Total number of sports balls collected by Mr. Trash Wheel in 2020
total_ball_mr = mr_df %>% 
  filter(year == "2020") %>% 
  pull(sports_balls) %>% 
  sum()

```

* There are `r nrow(merge_df)` observations in the merged dataset along with  `r ncol(merge_df)` variables. Those variables are colled  `r names(merge_df)`.
* The total weight of trash collected by Professor Trash Wheel is `r total_weight_prof` tons.
* The total number of sports balls collected by Mr. Trash Wheel in 2020 was `r total_ball_mr`.




### Problem 3

```{r}
# Tidy the pols_month dataset
pols_df = read.csv("./hw2_data/fivethirtyeight_datasets/pols-month.csv") |>  
  janitor::clean_names() |> 
  separate(mon, c("year","month","day")) |> 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) |> 
  mutate(month = month.abb[month]) |>  
  mutate(president = if_else(prez_dem == 1, "dem",
                     if_else(prez_gop > 0, "gop", ""))
           ) |>
  dplyr::select(-prez_dem, -prez_gop, -day)
```


```{r}

# Tidy the snp dataset
# ??? lubricate and year limitation 
snp_df = read.csv("./hw2_data/fivethirtyeight_datasets/snp.csv") |>  
  janitor::clean_names() |> 
  mutate(date = lubridate::parse_date_time2(date,orders ="mdy", cutoff_2000 = 23)) |>  
  separate(date, c("year","month","day")) |> 
  mutate(year = as.integer(year), month = as.integer(month), day = as.integer(day)) |> 
  mutate(month = month.abb[month]) |> 
  relocate(year, .before = "month") |>  
  select(-day)
```


```{r}
# Tidy the unemployment dataset
unemployment_df = read.csv("./hw2_data/fivethirtyeight_datasets/unemployment.csv") |>  
  pivot_longer(Jan:Dec,names_to = "month", values_to = "unemployment_rate") |> 
  janitor::clean_names() 

```

```{r}
# Merge snp into pols
snp_into_pols =  left_join(pols_df, snp_df, by = c("year","month"))
```


```{r}
# Merge unemployment into the result
merge_df = left_join(snp_into_pols, unemployment_df, by = c("year","month"))

```

 Datasets were taken from `FiveThirtyEight` data.

* The `pols-month` dataset had `r names(pols_df)` variables and ` r nrow(pols_df)` observation. Variable names as following  `r names(pols_df)`. Year range was (`r pols_df |>  pull(year) |> range()`). 

* The `snp` dataset had `r names(snp_df)` variables and ` r nrow(snp_df)` observation. Variable names as following  `r names(snp_df)`. Year range was (`r snp_df |>  pull(year) |> range()`). 

* The `unemployment` dataset had `r names(unemployment_df)` variables and ` r nrow(unemployment_df)` observation. Variable names as following  `r names(unemployment_df)`. Year range was (`r unemployment_df |>  pull(year) |> range()`).

* After merging `snp` dataset into `pols-month` dataset, the dimension of the resulting dataset is `r nrow(snp_into_pols)` rows x `r ncol(snp_into_pols)` columns. It includes `r ncol(snp_into_pols)` variables: `r names(snp_into_pols)`. The range of years is [`r snp_into_pols %>% pull(year) %>% range()`].

* After merging `unemployment` dataset into the previous resulting dataset `snp_into_pols`, the dimension of the merged dataset is `r nrow(merge_df)` rows x `r ncol(merge_df)` columns. It includes `r ncol(merge_df)` variables: `r names(merge_df)`. The range of years is [`r merge_df %>% pull(year) %>% range()`].





















The total number is `r prof_df %>% pull(weight) %>% sum()`